Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-19 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
added ref, mass flow, number of photos

@@

Table~\ref{tab:Flame_Heights} provides the mean visible flame height, as well as the range of flame heights for each fuel, based on analysis of the photographs for three replicate experiments.  In each case, the photographs were chosen for analysis while the fire was generating its peak heat release rate.  The range of flame heights for each fuel is approximately twice the minimum flame height or the continuous flame height. Analyzing the photographs in this manner was very time consuming and was based on making measurements on each photograph with a scale.  The limit of 5 frames per second also introduces additional uncertainty in determining the flame height. These values will serve as a basis for comparing with mean flame height values determined with video analysis.

@@

There have been a number of studies that have examined gypsum wallboard under elevated temperature conditions in terms of it's thermal properties or in terms of its physical change as a result to exposure to heat.  Schroeder conducted a study on the post-fire analysis of construction materials including gypsum.  As gypsum is heated to temperatures above 80~$^\circ$C it begins to change phase and degrade.  Continued heating causes some of the chemically bound water to disassociate and the gypsum changes to a hemihydrate.  This transformation occurs in the range of 98~$^\circ$C to 130~$^\circ$C. With additional heating the residue of the gypsum continues to transform and soften turning into anhydrite, commonly this process is known as calcination.  Schroder exposed samples of gypsum to a range of heat fluxes and temperatures to determine the time for 80~$^\circ$C (dihydrate phase), 200~$^\circ$C (hemihydrate phase), and 500~$^\circ$C (anhydrous phase) isotherms to penetrate to different depths in the gypsum.  The crystal phase and the depth of penetration of the phase change was determined with the use of X~\-ray diffraction  add Schroder.  NFPA 921 also discusses calcination and how the investigator can use the visual appearance or probing of the soft dyhydrated gypsum to map fire patterns.

>>>>

\subsubsection{Natural Gas Flow}

The natural gas flow to the burner is measured in terms of volumetric flow with a dry test meter, that is corrected for temperature and pressure. The flow rate was determined to be 4.5 cubic feet per minute as output from the meter or 2.1~liters per second.  This yields a mass flow of natural gas of approximately 2~g/s.

@@

Table~\ref{tab:Flame_Heights} provides the mean visible flame height, as well as the range of flame heights for each fuel, based on analysis of the photographs for three replicate experiments.  In each case, the photographs were chosen for analysis while the fire was generating its peak heat release rate.  The range of flame heights for each fuel is approximately twice the minimum flame height or the continuous flame height. Analyzing the photographs in this manner was very time consuming and was based on making measurements on each photograph with a scale.  Between 6 and 10 seconds of still images, or 30 to 50 frames, were used to approximate the flames heights provided in Table~\ref{tab:Flame_Heights}.  The limit of 5 frames per second also introduces additional uncertainty in determining the flame height. These values will serve as a basis for comparing with mean flame height values determined with video analysis.

@@

There have been a number of studies that have examined gypsum wallboard under elevated temperature conditions in terms of it's thermal properties or in terms of its physical change as a result to exposure to heat.  Schroeder conducted a study on the post-fire analysis of construction materials including gypsum.  As gypsum is heated to temperatures above 80~$^\circ$C it begins to change phase and degrade.  Continued heating causes some of the chemically bound water to disassociate and the gypsum changes to a hemihydrate.  This transformation occurs in the range of 98~$^\circ$C to 130~$^\circ$C. With additional heating the residue of the gypsum continues to transform and soften turning into anhydrite, commonly this process is known as calcination.  Schroder exposed samples of gypsum to a range of heat fluxes and temperatures to determine the time for 80~$^\circ$C (dihydrate phase), 200~$^\circ$C (hemihydrate phase), and 500~$^\circ$C (anhydrous phase) isotherms to penetrate to different depths in the gypsum.  The crystal phase and the depth of penetration of the phase change was determined with the use of X~\-ray diffraction~\cite{Schroeder:1999}.  NFPA 921 also discusses calcination and how the investigator can use the visual appearance or probing of the soft dihydrated gypsum to map fire patterns.
